Validation of an Actuator Line Model Coupled to a Dynamic Stall Model for Pitching Motions Characteristic to Vertical Axis Turbines

### 2016-08-21 paper.tex
Made by --> by
More instances of 'pithing'

Static lift and drag coefficients of the 2D profile used for this study are
taken from the technical report made by Sheldahl and Klimas
\cite{sheldahl1981aerodynamic}, which is a popular database containing values
for a wide range of Reynolds numbers. These coefficients are tabulated and
combined with the blade element approach for determining body forces acting on
blades. Relative flow velocity $\vec{V}_{\mathrm{rel}}$ and angle of attack
$\alpha$ are calculated for each blade using the geometric relation between the
velocity of the blade $- \Omega r$, with $\Omega$ representing the angular
(rotational) velocity of the rotor and $r$ the radius of the blade element, and
the velocity of the local incident wind flow $\vec{V}_{\mathrm{in}}$ which is generally lower than the incoming flow velocity $\vec{V}_{\infty}$. Figure
\ref{figvectors} depicts a cross-sectional airfoil element at radius $r$ in the
plane perpendicular to the turbine axis.

@@

The unsteady attached flow solution considers the circulatory and impulsive loads, which, at the same time depend on the unsteady bound vortex. The circulatory normal force coefficient could be expressed as
\begin{align}
& C_{N_n}^C = C_{N_\alpha} \alpha_{E_n}	\label{CNcirc}
\end{align}
where $C_{N \alpha}$ indicates the slope of the static normal force coefficient for a specific Reynolds number, and $\alpha_{E_n}$ is an expression for an equivalent angle of attack
\begin{align}
& \alpha_{E_n} = \alpha_n - X_n -Y_n - Z_n	\label{alphaeq}
\end{align}
where $\alpha$ is the geometrical angle of attack and $X$, $Y$ and $Z$ are the deficiency functions, which are empirically derived, based on the flow velocity and the pithing rate, the details are found in \cite{dyachuk2013dynamic}. Indices $n$ and $n-1$ represent the current and previous time-step. A delayed angle of attack is considered due to the lag in pressure response, and it is calculated as
\begin{align}
& {\alpha _n}' = {\alpha _n} - D_{\alpha _n}    	\label{alphadel}
\end{align}
with $D_\alpha$ as the deficiency function
\begin{align}
& D_{\alpha_n} = D_{\alpha_{n-1}} \exp \left( - \frac{\Delta s}{T_\alpha} \right) + (\alpha_n - \alpha_{n-1})\exp \left( - \frac{\Delta s}{2T_\alpha} \right) \label{defalpha}
\end{align}
with the empirically derived time constant $T_\alpha$, which has a value of $T_\alpha$ = 6.3 for the NACA0021 airfoil, and $\Delta s$ corresponds to a non-dimensional time-step
\begin{align}
& \Delta s = \frac{2 | \vec{V}_{rel} | \Delta t}{c}		.	\label{deltas}
\end{align}

@@

\begin{figure}[h]
\begin{minipage}{18pc}
\resizebox{\columnwidth}{!}{\input{./figures/CN226dtLEScoarse}}
%\resizebox{\columnwidth}{!}{\input{plot}}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
\resizebox{\columnwidth}{!}{\input{./figures/CT226dtLEScoarse}}
\end{minipage}
\caption{\label{figdtcoarse}Normal (left) and tangential (right) force coefficients varying time-step discretization for a pithing motion with a maximum amplitude of 22.6\degree\ }
\end{figure}

@@

It has been demonstrated that the presented coupled model ALM with DSM, shows
good agreement with experimental data, considering the complexity of the flows
studied. Moreover, the model has been tested for a wide range of pithing
amplitudes, therefore, with different stall conditions (none, shallow, and deep
stall), results agreed well compared with experiments. The model is stable and
accurate for the tested pithing blade motions, which makes it suitable for
application in VAWT simulations.

>>>>

Static lift and drag coefficients of the 2D profile used for this study are
taken from the technical report by Sheldahl and Klimas
\cite{sheldahl1981aerodynamic}, which is a popular database containing values
for a wide range of Reynolds numbers. These coefficients are tabulated and
combined with the blade element approach for determining body forces acting on
blades. Relative flow velocity $\vec{V}_{\mathrm{rel}}$ and angle of attack
$\alpha$ are calculated for each blade using the geometric relation between the
velocity of the blade $- \Omega r$, with $\Omega$ representing the angular
(rotational) velocity of the rotor and $r$ the radius of the blade element, and
the velocity of the local incident wind flow $\vec{V}_{\mathrm{in}}$ which is generally lower than the incoming flow velocity $\vec{V}_{\infty}$. Figure
\ref{figvectors} depicts a cross-sectional airfoil element at radius $r$ in the
plane perpendicular to the turbine axis.

@@

The unsteady attached flow solution considers the circulatory and impulsive loads, which, at the same time depend on the unsteady bound vortex. The circulatory normal force coefficient could be expressed as
\begin{align}
& C_{N_n}^C = C_{N_\alpha} \alpha_{E_n}	\label{CNcirc}
\end{align}
where $C_{N \alpha}$ indicates the slope of the static normal force coefficient for a specific Reynolds number, and $\alpha_{E_n}$ is an expression for an equivalent angle of attack
\begin{align}
& \alpha_{E_n} = \alpha_n - X_n -Y_n - Z_n	\label{alphaeq}
\end{align}
where $\alpha$ is the geometrical angle of attack and $X$, $Y$ and $Z$ are the deficiency functions, which are empirically derived, based on the flow velocity and the pitching rate, the details are found in \cite{dyachuk2013dynamic}. Indices $n$ and $n-1$ represent the current and previous time-step. A delayed angle of attack is considered due to the lag in pressure response, and it is calculated as
\begin{align}
& {\alpha _n}' = {\alpha _n} - D_{\alpha _n}    	\label{alphadel}
\end{align}
with $D_\alpha$ as the deficiency function
\begin{align}
& D_{\alpha_n} = D_{\alpha_{n-1}} \exp \left( - \frac{\Delta s}{T_\alpha} \right) + (\alpha_n - \alpha_{n-1})\exp \left( - \frac{\Delta s}{2T_\alpha} \right) \label{defalpha}
\end{align}
with the empirically derived time constant $T_\alpha$, which has a value of $T_\alpha$ = 6.3 for the NACA0021 airfoil, and $\Delta s$ corresponds to a non-dimensional time-step
\begin{align}
& \Delta s = \frac{2 | \vec{V}_{rel} | \Delta t}{c}		.	\label{deltas}
\end{align}

@@

\begin{figure}[h]
\begin{minipage}{18pc}
\resizebox{\columnwidth}{!}{\input{./figures/CN226dtLEScoarse}}
%\resizebox{\columnwidth}{!}{\input{plot}}
\end{minipage}\hspace{2pc}%
\begin{minipage}{18pc}
\resizebox{\columnwidth}{!}{\input{./figures/CT226dtLEScoarse}}
\end{minipage}
\caption{\label{figdtcoarse}Normal (left) and tangential (right) force coefficients varying time-step discretization for a pitching motion with a maximum amplitude of 22.6\degree\ }
\end{figure}

@@

It has been demonstrated that the presented coupled model ALM with DSM, shows
good agreement with experimental data, considering the complexity of the flows
studied. Moreover, the model has been tested for a wide range of pitching
amplitudes, therefore, with different stall conditions (none, shallow, and deep
stall), results agreed well compared with experiments. The model is stable and
accurate for the tested pitching blade motions, which makes it suitable for
application in VAWT simulations.
